Apparent preferences for cognitive effort fade when multiple forms of effort

### 2021-05-29 Manuscript.Rmd
into the discussion edits

  Understanding the sporadic attractiveness of effort remains elusive. In this study, we sought to identify factors underlying this phenomenon, and tested the hypothesis that the perceived cost or value of a demand could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found a tendency toward engaging with cognitively effortful prospets when individuals faced a single form of demand, which faded over time when other costs were also experienced.

  Our findings provided partial support for our pre-registered hypotheses (https://osf.io/2rsgm/registrations). In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal points. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [@Constantino2015; @Wikenheiser2013]. In contrast, our prediction that effort requirements would produce more acceptances only held for the cognitive group, and this tendency for higher acceptances did not switch to passive waiting as predicted for Experiment 2. These hypotheses originally stemmed from prevalent proposals that incentivized work should become more appealing [@Eisenberger1992; @Inzlicht2018], but that people would opt for the easiest available way to achieve desired outcomes [@Hull1943; @Olivola2013]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted that reward-independent task features can promote the pursuit of effortful prospects. The fact that predicted stable preferences for cognitive effort carried over onto the beginning of Experiment 2 further supported a task-specific modulation. However, the unexpected gradual convergence in acceptances in the second experiment emphasized that these features could be contextually molded.

>>>>

  Understanding the sporadic attractiveness of effort remains elusive. In this study, we sought to identify factors underlying this phenomenon, and tested the hypothesis that the perceived cost or value of effort could depend on whether the decision environment contains opportunities to achieve the same outcome through different means. We conducted two foraging experiments that exposed individuals to a single or intermixed types of demands (cognitive effort, physical effort, or passive delay). We found a tendency toward engaging with cognitively effortful prospects when individuals faced a single form of demand, which faded over time when other costs were also experienced.

  Our findings provided partial support for our pre-registered hypotheses (https://osf.io/2rsgm/registrations). In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal levels. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [@Constantino2015; @Wikenheiser2013]. In contrast, our prediction that effort requirements would lead to greater acceptance rates only held for the cognitive effort group, and this tendency did not reverse as predicted in Experiment 2. These hypotheses originally stemmed from prevalent proposals that incentivized work should become more appealing when presented in isolation [@Eisenberger1992; @Inzlicht2018], but that people would opt for the easiest way to achieve desired outcomes when multiple alternatives are available [@Hull1943; @Kool2010; @Olivola2013]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted that reward-independent task characteristics can promote or inhibit the pursuit of effortful prospects.
